---
documentclass: article
fontsize: 12pt
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: 
    toc: false
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
      before_body: titlepage.sty
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rprojroot)
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r load, include=FALSE, message=FALSE, warning=FALSE}
load(find_root_file("data/cosmetics.Rdata", 
     criterion = has_file("MultivariateStatistics_assignment.Rproj")))
```

## CFA to construct a measurement model for the Attitude items
There are `r sum(str_detect(colnames(cosmetics),"Attitude"))` attitude items that are scored on a five-point Likert scale. 

## CFA to construct a measurement model for the Behavior-Intention items
There are `r sum(str_detect(colnames(cosmetics),"BI"))` attitude items that are scored on a five-point Likert scale. 

## Structural equation model to evaluate the impact of attitude on behavior intention

# Task 2

```{r load2, include=FALSE, message=FALSE, warning=FALSE}
load(find_root_file("data/benefits.Rdata", 
     criterion = has_file("MultivariateStatistics_assignment.Rproj")))
```

## Canonical correlation analysis


## Split-half approach
