---
documentclass: article
fontsize: 10pt
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: 
    fig_caption: yes
    toc: false
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
      before_body: titlepage.sty
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}

update.packages("rlang")
library(xfun)
library(tidyverse)
library(rprojroot)
library(kableExtra)
library(MASS)#lda function
library(biotools)
library(ISLR2)
# this package does not work for the newest version of R --> 
# only load if the local data files are not there
if (!file.exists(find_root_file("data/rect_constr.csv", 
     criterion = has_file("MultivariateStatistics_assignment.Rproj")))) {
  library(smacof)
} 
 
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

All R scripts and the data can be found on [this GitHub repository](https://github.com/raiisac/MultivariateStatistics_assignment).
```{r load, include=FALSE, message=FALSE, warning=FALSE}
data(College)

#test the difference between centroids
Wilks_manova <- function(data, group = "Private"){
  lm.out <- lm(cbind(
    sapply(X = colnames(data)[colnames(data) != group], 
           FUN = function(x) get(x)))~ 
       get((group)), 
   data = data)
  s <- summary(manova(lm.out), test = c("Wilks"))
  return(s$stats[1, "Pr(>F)"])#return the pvalue
}

boxM_test <- function(data, group = "Private"){
  out <- boxM(data[,colnames(data) != group], data[,group])
  return(out$p.value)#return the pvalue
}

lda_analysis <- function(data, group = "Private"){
  lda.out <- lda(
    formula(paste0(group, "~", paste(colnames(data)[colnames(data) != group], 
                                     collapse = "+"))),
    data = data)
}

LCA <- function(data, group="Private")

```

The data consists of `r ncol(College)` variables of which the categorical variable *Private* denotes whether the school is private (*Private == Yes*) or not (*Private == No*). There are `r nrow(College)` schools in the data; `r sum(College$Private == "Yes")` (`r round(sum(College$Private == "Yes")/nrow(College)*100,2)`%) private schools and `r sum(College$Private == "No")` (`r round(sum(College$Private == "No")/nrow(College)*100,2)`%) non-private schools.

Before we start to build models to distinguish private and non-private schools, Wilk's lambda test is performed where the null hypothesis states that the centroid (mean) in both groups is the same. Both in the log-transformed and the original (centered or standardized) data, this null hypothesis is rejected. This means that the centroids of private and non-private schools differ significantly

```{r classification, include=FALSE, message=FALSE, warning=FALSE}
College_logtranfskewed <- College %>%
  mutate(Apps = log(Apps),
         Accept = log(Accept),
         Enroll = log(Enroll),
         Top10perc = log(Top10perc),
         F.Undergrad = log(F.Undergrad),
         P.Undergrad = log(P.Undergrad),
         Books = log(Books),
         Personal = log(Personal),
         Expend = log(Expend))
College_centered <- College %>% 
  dplyr::select(-Private) %>% 
  scale(center = TRUE, scale = FALSE) %>%
  as.data.frame() %>%
  mutate(Private = College$Private)
College_logtranfskewed_centered <- College_logtranfskewed %>% 
  dplyr::select(-Private) %>%  
  scale(center = TRUE, scale = FALSE) %>%
  as.data.frame() %>%
  mutate(Private = College$Private)
College_std <- College %>% 
  dplyr::select(-Private) %>%  
  scale(center = TRUE, scale = TRUE) %>%
  as.data.frame() %>%
  mutate(Private = College$Private)
College_logtranfskewed_std <- College_logtranfskewed %>% 
  dplyr::select(-Private) %>%  
  scale(center = TRUE, scale = TRUE) %>%
  as.data.frame() %>%
  mutate(Private = College$Private)


#test the difference between centroids
Wilks_manova(data = College_logtranfskewed_centered, group = "Private")
Wilks_manova(data = College_centered, group = "Private")
Wilks_manova(data = College_logtranfskewed_std, group = "Private")
Wilks_manova(data = College_std, group = "Private")
#In each dataset, H0:mu_{yes} = mu_{no} is rejected --> centroids of Private and non-private schools differ significantly

#LDA

#QDA

#bagging

#random forsts

#KNN

```


# Task 2

```{r load2, include=FALSE, message=FALSE, warning=FALSE}
load(find_root_file("data/fashion.Rdata", 
     criterion = has_file("MultivariateStatistics_assignment.Rproj")))
train.data <- train.data %>% scale(center = TRUE, scale = FALSE)
```

# Task 3
```{r load3, include=FALSE, message=FALSE, warning=FALSE}
if (file.exists(find_root_file("data/rect_constr.csv", 
     criterion = has_file("MultivariateStatistics_assignment.Rproj")))) {
  rect_constr <- read.csv(
    find_root_file("data/rect_constr.csv", 
                   criterion = 
                     has_file("MultivariateStatistics_assignment.Rproj")))
} else {
  data(rect_constr) 
}
if (file.exists(find_root_file("data/rectangles.csv", 
     criterion = has_file("MultivariateStatistics_assignment.Rproj")))) {
  rectangles <- read.csv(
    find_root_file("data/rectangles.csv", 
                   criterion = 
                     has_file("MultivariateStatistics_assignment.Rproj")))
} else {
  data(rectangles) 
}
```

# Appendix

